---
title: "ma615_midterm project"
author: "xiaofan xia"
date: "2019/10/15"
output: 
  ioslides_presentation:
    incremental: true
    
---

```{r setup, include=FALSE}
library(tidyverse)
```


## The dataset is from the world bank open data. This is a dataset of the health nutrition and population. It contains 259 countries and 59 years with 10 series of variables.


## Firstly, I read in the data and clean the data.I specifically choose the variable of adolescent fertility rate (birth per 1,000 women ages 15-19). 

```{r warning=FALSE, echo=FALSE}
data <- read.csv("data2.csv",encoding = "UTF-8") 
year <- colnames(data)[5:63]
data_new <- 
  data %>%
  gather(year,key = "year",value = "value")%>%
  select(Series.Name,Country.Name,year,value)
data_new$year <- as.numeric(str_sub(data_new$year,2,5))
data_new$value <- as.numeric(data_new$value) 

data_adolscent_fert_rate <-
  data_new%>%
  filter(Series.Name=="Adolescent fertility rate (births per 1,000 women ages 15-19)")
```


## I did EDA by world areas. First one is East Asia and Pacific.
```{r warning=FALSE,echo=FALSE}
East_Asia_Pacific <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="East Asia & Pacific")
ggplot(East_Asia_Pacific)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of East Asia & Pacific")
```

## Arab World
```{r warning=FALSE,echo=FALSE}
Arab_World <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Arab World")
ggplot(Arab_World)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of Arab World")
```

## Europe and Central Asia
```{r warning=FALSE,echo=FALSE}
Europe_Central_Asia <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Europe & Central Asia")
ggplot(Europe_Central_Asia)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of Europe & Central Asia")
```

## Latin America and Caribbean
```{r warning=FALSE,echo=FALSE}
Latin_America_Caribbean <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Latin America & Caribbean")
ggplot(Latin_America_Caribbean)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of Latin America & Caribbean")
```

## Middle East and North Africa
```{r warning=FALSE,echo=FALSE}
Middle_East_North_Africa <-
 data_adolscent_fert_rate%>%
  filter(Country.Name=="Middle East & North Africa")
ggplot(Middle_East_North_Africa)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of Middle East & North Africa")
```

## North America
```{r warning=FALSE,echo=FALSE}
North_America <-
   data_adolscent_fert_rate%>%
  filter(Country.Name=="North America")
ggplot(North_America)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of North America")
```

## South Asia
```{r warning=FALSE, echo=FALSE}
South_Asia <-
   data_adolscent_fert_rate%>%
  filter(Country.Name=="South Asia")
ggplot(South_Asia)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of South Asia")
```

## Central Europe and the Baltics
```{r warning=FALSE, echo=FALSE}
Central_Europe_the_Baltics <-
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Central Europe and the Baltics")
ggplot(Central_Europe_the_Baltics)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of Central Europe and the Baltics")
```

## World
```{r warning=FALSE, echo=FALSE}
World <-
  data_adolscent_fert_rate%>%
  filter(Country.Name=="World")
ggplot(World)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of the World")
```

## Differences of different world areas
```{r warning=FALSE,echo=FALSE}
region <- rbind(World,Central_Europe_the_Baltics,South_Asia,North_America,Middle_East_North_Africa,Latin_America_Caribbean,Europe_Central_Asia,Arab_World,East_Asia_Pacific)

ggplot(region)+aes(year,value)+geom_smooth(aes(color=Country.Name))+ggtitle("Adolescent fertility rate for different world areas")+theme(legend.position = "bottom") #show differences regarding to different world areas
```

## EDA by income level- Low Income Countries
```{r warning=FALSE,echo=FALSE}
low_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Low income")
ggplot(low_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of low income countries")

```

## Low and Middle Income Countries
```{r warning=FALSE,echo=FALSE}
low_middle_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Low & middle income")
ggplot(low_middle_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of low and middle income countries")
```


## Lower Middle Income Countries
```{r warning=FALSE,echo=FALSE}
lower_middle_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Lower middle income")
ggplot(lower_middle_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of lower middle income countries")
```

## Middle Income Countries
```{r warning=FALSE,echo=FALSE}
middle_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Middle income")
ggplot(middle_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of middle income countries")
```

## Upper Income Countries
```{r warning=FALSE,echo=FALSE}
upper_middle_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="Upper middle income")
ggplot(upper_middle_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of upper middle income countries")
```

## High Income Countries
```{r warning=FALSE,echo=FALSE}
high_income <- 
  data_adolscent_fert_rate%>%
  filter(Country.Name=="High income")
ggplot(high_income)+aes(year,value)+geom_point(aes(color=year))+ggtitle("Adolescent fertility rate of high income countries")
```

## Income level differences 
```{r warning=FALSE,echo=FALSE}
income <-
  data_adolscent_fert_rate%>%
  filter(grepl("income",data_adolscent_fert_rate$Country.Name)==TRUE)
ggplot(income)+aes(year,value)+geom_smooth(aes(color=Country.Name))+ggtitle("Adolescent fertility rate regarding to different income level")+theme(legend.position = "bottom") #compare the rate regarding to different income levels
```
